# Analyses result tables {#analyses-tables}

<!-- TODO: add links to all tables (e.h. to OMIM, PubMed, PanelApp etc.) -->

## Main table: Merged analyses sources

This table shows the merged results of all analyses files as a wide table with summarized information.

```{r, echo=FALSE}
merge_analyses_sources_reformat <- merge_analyses_sources %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol)

datatable(merge_analyses_sources_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE, # this argument renders the links as HTML
          extensions = 'Buttons', # this argument adds an extension for download buttons
          options = list(
            dom = 'Blfrtip',
            scrollX = '400px',
            scroller = TRUE,
            buttons = c('copy',
              'csv',
              'excel',
              'pdf'), # here we add the download buttons for different formats
            lengthMenu = list(c(10, 30, 50, -1), 
                              c('10', '30', '50', 'All')) # here we change the selection options
          )
        )
```


## Result table: PanelApp

This table shows results of the first analysis searching kidney disease associated genes from the PanelApp project in the UK and Australia.

```{r, echo=FALSE}
PanelApp_genes_reformat <- PanelApp_genes %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol, -gene_name_reported)

datatable(PanelApp_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: Literature

This table shows results of the second analysis searching kidney disease associated genes from various publications.

```{r, echo=FALSE}
Literature_genes_reformat <- Literature_genes %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol, -gene_name_reported)

datatable(Literature_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: Diagnostic panels

This table shows results of the third analysis searching kidney disease associated genes from clinical diagnostic panels for kidney disease.

```{r, echo=FALSE}
DiagnosticPanels_genes_reformat <- DiagnosticPanels_genes %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol, -gene_name_reported)

datatable(DiagnosticPanels_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: HPO in rare disease databases

This table shows results of the fourth analysis searching kidney disease associated genes from a Human Phenotype Ontology (HPO)-based search in rare disease databases (OMIM, Orphanet).

```{r, echo=FALSE}
HPO_genes_reformat <- HPO_genes %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol, -gene_name_reported)

datatable(HPO_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: PubTator

This table shows results of the fifth analysis searching kidney disease associated genes from a PubTator API-based automated literature extraction from PubMed.

```{r, echo=FALSE}
PubTator_genes_reformat <- PubTator_genes %>%
  mutate(gene = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" ><i>',
    approved_symbol,
    '</i></a>'),
    , .before = approved_symbol) %>%
  select(-hgnc_id, -approved_symbol, -gene_name_reported)

datatable(PubTator_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```