# Analyses result tables {#analyses-tables}

<!-- here we load required libraries -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Load libraries
library(readr) # needed to import CSV result files
library(tidyverse) # needed to transform the tables
library(DT) # needed to generate nice data tables in html
library(plotly) # needed for upset plots
```

<!-- here we load the result csv files -->
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## load the CSV data
# TODO: load and filter (newest) the files automatically 
merge_analyses_sources <- read_csv("../analyses/A_MergeAnalysesSources/A_MergeAnalysesSources.2023-05-18.csv.gz")
PanelApp_genes <- read_csv("../analyses/01_PanelApp/results/01_PanelApp_genes.2023-05-18.csv.gz")
Literature_genes <- read_csv("../analyses/02_Literature/results/02_Literature_genes.2023-05-18.csv.gz")
DiagnosticPanels_genes <- read_csv("../analyses/03_DiagnosticPanels/results/03_DiagnosticPanels_genes.2023-05-18.csv.gz")
HPO_genes <- read_csv("../analyses/04_HPO/results/04_HPO_genes.2023-05-18.csv.gz")
PubTator_genes <- read_csv("../analyses/05_PubTator/results/05_PubTator_genes.2023-05-18.csv.gz")
```
---


## Main table: Merged analyses sources

This table shows the merged results of all analyses files as a wide table with summarized information.

```{r, echo=FALSE}
merge_analyses_sources_reformat <- merge_analyses_sources %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(merge_analyses_sources_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE, # this argument renders the links as HTML
          extensions = 'Buttons', # this argument adds an extension for download buttons
          options = list(
            dom = 'Blfrtip',
            buttons = c('copy',
              'csv',
              'excel',
              'pdf'), # here we add the download buttons for different formats
            lengthMenu = list(c(10, 30, 50, -1), 
                              c('10', '30', '50', 'All')) # here we change the selection options
          )
        )
```


## Result table: PanelApp

This table shows results of the first analysis searching kidney disease associated genes from the PanelApp project in the UK and Australia.

```{r, echo=FALSE}
PanelApp_genes_reformat <- PanelApp_genes %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(PanelApp_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: Literature

This table shows results of the second analysis searching kidney disease associated genes from various publications.

```{r, echo=FALSE}
Literature_genes_reformat <- Literature_genes %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(Literature_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: Diagnostic panels

This table shows results of the third analysis searching kidney disease associated genes from clinical diagnostic panels for kidney disease.

```{r, echo=FALSE}
DiagnosticPanels_genes_reformat <- DiagnosticPanels_genes %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(DiagnosticPanels_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: HPO in rare disease databases

This table shows results of the fourth analysis searching kidney disease associated genes from a Human Phenotype Ontology (HPO)-based search in rare disease databases (OMIM, Orphanet).

```{r, echo=FALSE}
HPO_genes_reformat <- HPO_genes %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(HPO_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```


## Result table: PubTator

This table shows results of the fifth analysis searching kidney disease associated genes from a PubTator API-based automated literature extraction from PubMed.

```{r, echo=FALSE}
PubTator_genes_reformat <- PubTator_genes %>%
  mutate(hgnc_id = paste0('<a href="https://www.genenames.org/data/gene-symbol-report/#!/hgnc_id/HGNC:',
    hgnc_id,
    '" target="_blank" >',
    hgnc_id,
    '</a>')) %>%
  mutate(approved_symbol = paste0('<i>',
    approved_symbol,
    '</i>'))

datatable(PubTator_genes_reformat,
          filter = 'top', # this argument positions the filtering input at the top of the table
          escape = FALSE) # this argument renders the links as HTML
```