# Analyses plots {#analyses-plots}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## load the CSV data
# TODO: load and filter (newest) the files automatically 
merge_analyses_sources <- read_csv("../analyses/merged/KidneyGenetics_MergeAnalysesSources.2023-05-18.csv")
PanelApp_genes <- read_csv("../analyses/01_PanelApp/results/01_PanelApp_genes.2023-05-18.csv")
Literature_genes <- read_csv("../analyses/02_Literature/results/02_Literature_genes.2023-05-18.csv")
DiagnosticPanels_genes <- read_csv("../analyses/03_DiagnosticPanels/results/03_DiagnosticPanels_genes.2023-05-18.csv")
HPO_genes <- read_csv("../analyses/04_HPO/results/04_HPO_genes.2023-05-18.csv")
PubTator_genes <- read_csv("../analyses/05_PubTator/results/05_PubTator_genes.2023-05-18.csv")
```
---


## UpSet plot of merged analyses sources

UpSet plot of the merged analyses.

```{r, echo=FALSE, warning=FALSE}
analyses <- colnames(merge_analyses_sources)[5:9]

ComplexUpset::upset(merge_analyses_sources, analyses)
```


## Bar plot of PanelApp results

Bar plot of the PanelApp analysis.

```{r, echo=FALSE, warning=FALSE}
# compute counts
PanelApp_genes_counts <- PanelApp_genes %>%
  group_by(source_count) %>%
  summarise(number = n())

# generate plot
PanelApp_genes_counts_plot <- ggplot(PanelApp_genes_counts , aes(x = source_count, y = number)) + 
    geom_bar(stat = "identity") +
    theme_classic()

# convert to interactive using ggplotly
ggplotly(PanelApp_genes_counts_plot)
```


## Bar plot of Literature results

Bar plot of the Literature analysis.

```{r, echo=FALSE, warning=FALSE}
# compute counts
Literature_genes_counts <- Literature_genes %>%
  group_by(source_count) %>%
  summarise(number = n())

# generate plot
Literature_genes_counts_plot <- ggplot(Literature_genes_counts , aes(x = source_count, y = number)) + 
    geom_bar(stat = "identity") +
    theme_classic()

# convert to interactive using ggplotly
ggplotly(Literature_genes_counts_plot)
```


## Bar plot of Diagnostic panels results

Bar plot of the Diagnostic panels analysis.

```{r, echo=FALSE, warning=FALSE}
# compute counts
DiagnosticPanels_genes_counts <- DiagnosticPanels_genes %>%
  group_by(source_count) %>%
  summarise(number = n())

# generate plot
DiagnosticPanels_genes_counts_plot <- ggplot(DiagnosticPanels_genes_counts , aes(x = source_count, y = number)) + 
    geom_bar(stat = "identity") +
    theme_classic()

# convert to interactive using ggplotly
ggplotly(DiagnosticPanels_genes_counts_plot)
```


## Bar plot of HPO in rare disease databases results

Bar plot of the HPO in rare disease databases analysis.

```{r, echo=FALSE, warning=FALSE}
# compute counts
HPO_genes_counts <- HPO_genes %>%
  group_by(source_count) %>%
  summarise(number = n())

# generate plot
HPO_genes_counts_plot <- ggplot(HPO_genes_counts , aes(x = source_count, y = number)) + 
    geom_bar(stat = "identity") +
    theme_classic()

# convert to interactive using ggplotly
ggplotly(HPO_genes_counts_plot)
```


## Bar plot of PubTator results

Bar plot of thePubTator analysis.

```{r, echo=FALSE, warning=FALSE}
# compute counts
PubTator_genes_counts <- PubTator_genes %>%
  group_by(source_count) %>%
  summarise(number = n())

# generate plot
PubTator_genes_counts_plot <- ggplot(PubTator_genes_counts , aes(x = source_count, y = number)) + 
    geom_bar(stat = "identity") +
    theme_classic()

# convert to interactive using ggplotly
ggplotly(PubTator_genes_counts_plot)
```