# Visualizations {#visualizations}

This section provides a comprehensive overview of presentations and posters related to the Kidney Genetics project. 
As an essential part of our approach to transparency, we have assembled a collection of materials,
presented at various conferences and meetings to illustrate the milestones in the project's development.


## Objectives

1. **Project Progress Tracking:** By organizing presentations and posters chronologically, this section serves as a dynamic timeline to track the progress and evolution of the project.

2. **Event-specific Insights:** Each entry includes details about the date and event where the presentation or poster was showcased.


## Presentations

> Below you can explore the list of Presentations:

**2023-12-05** / This is the presentation of our project presented at our monthly AG-Halbritter Working Group Meeting as a progress report.

```{r, echo=FALSE, warning=FALSE, out.width = "100%"}
iframe(width = "1600", height = "900", url_link = "static/img/visualizations/Kidney-Genetics_AG Halbritter_20231205.html")
```


**2023-10-07** / This is the presentation of our project presented by Nina Rank as a pitch talk at the annual congress of the German Society of Nephrology - DGfN in Berlin.

```{r, echo=FALSE, warning=FALSE, out.width = "100%"}
iframe(width = "1600", height = "900", url_link = "static/img/visualizations/Kidney-Genetics_DGfN_20231007.html")
```


**2023-07-05** / This is one of the first presentations at the beginning of this project with some initial ideas presented at our weekly AG-Halbritter Working Group Meeting.

```{r, echo=FALSE, warning=FALSE, out.width = "100%"}
iframe(width = "1600", height = "900", url_link = "static/img/visualizations/Kidney-Genetics_AG-Halbritter_20230705.html")
```



## Posters

> Below you can explore the list of Posters:

**2023-09-28** / This is our first Poster presented at the inaugural Symposium of the Max Rubner Center for Cardiovascular-Metabolic-Renal-Research.

```{r, echo=FALSE, warning=FALSE, out.width = "100%"}
iframe(width = "900", height = "1273", url_link = "static/img/visualizations/Kidney-Genetics_MRC Symposium_20230928.html")
```